Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest \\\large{A Bayesian Perspective}

### 2023-02-07 manuscript/manuscript.qmd
updated priors

```{r cache=TRUE}
# - INCEPTION alone default vague priors
set.seed(1234)
fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0)
# summary(fit1, digits=3)
# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
# get_variables(fit1)
# prior_summary(fit1)
# help('prior_summary.stanreg')

# informative priors using 2 previous studies
# one method
t1 <- rbeta(100000,25,124)
t2 <- rbeta(100000,45,95)
t3 <- t2 - t1
# t3 can be approximated by a normal distribution with the following parameters
# mean(t3)
# [1] 0.154
# sd(t3)
# [1] 0.0498

# second method
library(LearnBayes)
quantile1=list(p=.025, x=quantile(t3,0.025))     # 2.5% quantile should be 0.0556 
quantile2=list(p=.975, x=quantile(t3,0.975))      # 97.5% quantile should be 0..251
par_beta <- beta.select(quantile1, quantile2)
m_beta <-  par_beta[1]/(par_beta[1]+par_beta[2])
v_beta <- (par_beta[1] * par_beta[2])/((par_beta[1]+par_beta[2])^2 * (par_beta[1] +par_beta[2] + 1))
sd_beta <- sqrt(v_beta)
# Beta (6.35, 39.54) gives mean = 0.138 sd = 0.0504 close to estimates using method 1

# stan_glm doesn't seem able to take a Beta prior distribution so will use normal approximation

fit2 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0,
                 prior = normal(mean(t3), sd(t3)))
summary(fit2, digits=4)

**Have obviously not done this correctly as the prior should have shifted the default model parameter for eCPR to a larger number with a small sd**

@@

```{r}

# tt <- fit1 %>%
#  spread_draws(`(Intercept)`, `TxeCPR`)
# sum(tt$TxeCPR>0)/nrow(tt)

fit1 %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior") +
  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 71%") +
  theme_classic() +
  theme(legend.position="none")

# Examining prior information
x <- seq(0, .5, by = 0.001)

#THEN EACH OF OUR BETAS DISTR

beta1 <- dbeta(x, shape1 = 25, shape2 = 124) # probability of success with CPR
beta2 <- dbeta(x, shape1 = 45, shape2 = 95)  # probability of success with eCPR
beta3 <- dbeta(x, shape1 = 6.36, shape2 = 39.65)

mixture <- ggplot()+
  geom_line(aes(x, beta1),  color = "red",
            linewidth = 1.2) +
  geom_line(aes(x, beta2), color = "green",
            linewidth = 1.2) +
  geom_line(aes(x, beta3), color = "black",
            linewidth = 1.2) +
  labs(title = "Figure 2 Beta Distribution of prior studies",
       x = "Sequence",
       y = "Beta",
       caption = "Green = p(+outcome | eCPR) \nRed = p(+outcome | CPR) \nBlack= difference of the 2 probabilities")

mixture

>>>>

```{r cache=TRUE}
# - INCEPTION alone default vague priors
set.seed(1234)
fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
fit1_sum <- summary(fit1, digits=3)
conf_95 <- c(exp(fit1_sum[2]),exp(fit1_sum[2]+c(-1,1)*1.96*fit1_sum[10]))
tt <- fit1 %>%
    spread_draws(`(Intercept)`, `TxeCPR`)
fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
# summary(fit1, digits=3)
# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
# get_variables(fit1)
# prior_summary(fit1)
# help('prior_summary.stanreg')

#################
# function to calculate point estimate, 95% CrI, and P(HR>1)
# using prior information
################
eCRP_bayes <- function(data, m_CPR, sd_CPR, m_eCPR, sd_eCPR ) {
  set.seed(1234)
  fit <- stan_glm(prop_success ~ Tx, data = data, family = binomial(link="logit"), weights = total, refresh=0,
                  prior = normal(m_eCPR, sd_eCPR),
                  prior_intercept = normal (m_CPR, sd_CPR))
  fit_sum <- summary(fit, digits=3)
  conf_95 <- c(exp(fit_sum[2]),exp(fit_sum[2]+c(-1,1)*1.96*fit_sum[10]))
  tt <- fit %>%
    spread_draws(`(Intercept)`, `TxeCPR`)
  fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
  fit_draws <- fit %>% tidybayes::tidy_draws() 
  my_list <- list(fig_prob, conf_95, fit_draws)
  return(my_list)
}

# combined_prior 
# informative priors using 2 previous studies combined
comb_m_CPR <- sum(df1[c(2,3),4]) / (sum(df1[c(2,3),4]) + sum(df1[c(2,3),2])) # proportion previous results for CPR
comb_sd_CPR <- sqrt(comb_m_CPR*(1-comb_m_CPR))                               
comb_m_eCPR <- sum(df1[c(2,3),5]) / (sum(df1[c(2,3),5]) + sum(df1[c(2,3),3])) # proportion previous results for eCPR
comb_sd_eCPR <- sqrt(comb_m_eCPR*(1-comb_m_eCPR))
comb_prior <- eCRP_bayes(df[c(1:2),], comb_m_CPR, comb_sd_CPR, comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior

# enthusiastic prior with only ARREST study
enth_m_CPR <- sum(df1[c(2),4]) + 1 / (sum(df1[c(2),4]) + sum(df1[c(2),2])) # proportion previous results for CPR, + 0.5 to avoid 0 cell
enth_sd_CPR <- sqrt(enth_m_CPR*(1-enth_m_CPR))
enth_m_eCPR <- sum(df1[c(2),5]) / (sum(df1[c(2),5]) + sum(df1[c(2),3])) # proportion previous results for eCPR
enth_sd_eCPR <- sqrt(enth_m_eCPR*(1-enth_m_eCPR))
enth_prior <- eCRP_bayes(df[c(1:2),], enth_m_CPR, enth_sd_CPR, enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior

# skeptical prior with only PRAGUE study
skep_m_CPR <- sum(df1[c(3),4])  / (sum(df1[c(3),4]) + sum(df1[c(3),2])) # proportion previous results for CPR
skep_sd_CPR <- sqrt(skep_m_CPR*(1-skep_m_CPR))
skep_m_eCPR <- sum(df1[c(3),5]) / (sum(df1[c(3),5]) + sum(df1[c(3),3])) # proportion previous results for eCPR
skep_sd_eCPR <- sqrt(skep_m_eCPR*(1-skep_m_eCPR))
skep_prior <- eCRP_bayes(df[c(1:2),], skep_m_CPR, skep_sd_CPR, skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior

# skep_prior <- eCRP_bayes(df[c(1:2),], enth_m_CPR, enth_sd_CPR, skep_m_eCPR, skep_sd_eCPR) #test to see impact of enth intercept

table_tbl <- 
  tibble(
    prior = c("vague", "combined", "enthusiastic", "skeptical"),
    pt_est <- c(conf_95[1], comb_prior[[2]][1], enth_prior[[2]][1], skep_prior[[2]][1]),
    llimit <- c(conf_95[2], comb_prior[[2]][2], enth_prior[[2]][2], skep_prior[[2]][2]),
    ulimit <- c(conf_95[3], comb_prior[[2]][3], enth_prior[[2]][3], skep_prior[[2]][3]),
    p.gt.1 <- c(fig_prob, comb_prior[1], enth_prior[1], skep_prior[1])
  )

library(gt)
gt(table_tbl) %>%
  tab_stubhead(label = "Priors") 
```

# **Summary of what I don't understand**

This makes only partial sense. 95% CrI are narrower with informative priors. OK, but one would think the enthusiastic prior from a trial ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes for eCPR vs CPR would shift the HR of 1.32 to the right but instead the HR falls to have a 1.27. The skeptical prior from PRAGUE which was stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as expected

So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)    
skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)

Tx for enthusiastic prior > Tx skeptical prior (eCPR) but this seems to be offset by the reverse occurring for the intercept (CPR).     
The problem seems to be in how comparable are these previous RCTs as the baseline success for the standard TX (CRP) is radically different 0/14 vs 24/132.

When I use a prior that uses the enthusiastic intercept prior and the skeptical Tx prior, I  get a bigger downward shift with HR 1.220 95% CrI(0.678 2.194) and P>1 74.5%

Can you see any other explanation?

Do you have any other insights?

@@

```{r warning=FALSE, message=FALSE}

#############
# function to plot probability function
#############

plot_bayes <- function(data,text1,text2, fig_prob=.5){
  annon <- paste0(text2, fig_prob*100, "%")
  ggplot(data, aes(x = exp(TxeCPR),
                   fill = after_stat(x > 1),
                   slab_alpha = after_stat((x)))
  ) +
    ggdist::stat_halfeye() +
    scale_x_continuous(name=text, limits=c(0, 4)) +
    scale_y_continuous(name="Density") +
    labs(title = text1) +
    annotate(geom="text", x=2.75,y=.6, label = annon) +
    theme_classic() +
    theme(legend.position="none") 
}

#plot combined prior
#text1 <-c("HR increased survival with extracorporeal CPR")
#text2 <- c("Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR >")
#plot_bayes(comb_prior[[3]], text1, text2, comb_prior[[1]])

fit1 %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="HR being alive with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior") +
  annotate(geom="text", x=2.75,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7%") +
  theme_classic() +
  theme(legend.position="none")

### 2023-02-07 manuscript/manuscript.tex
updated priors

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
  pdfauthor={James M Brophy},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\begin{verbatim}

Attaching package: 'LearnBayes'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:rstanarm':

    laplace
\end{verbatim}

\begin{verbatim}

Model Info:
 function:     stan_glm
 family:       binomial [logit]
 formula:      prop_success ~ Tx
 algorithm:    sampling
 sample:       4000 (posterior sample size)
 priors:       see help('prior_summary')
 observations: 2
 predictors:   2

Estimates:
              mean    sd      10%     50%     90%  
(Intercept) -1.5987  0.2265 -1.8906 -1.5963 -1.3088
TxeCPR       0.1542  0.0501  0.0897  0.1535  0.2181

Fit Diagnostics:
           mean    sd      10%     50%     90%  
mean_PPD 11.9740  3.1650  8.0000 12.0000 16.0000

The mean_ppd is the sample average posterior predictive distribution of the outcome variable (for details see help('summary.stanreg')).

MCMC diagnostics
              mcse   Rhat   n_eff
(Intercept)   0.0042 0.9995 2872 
TxeCPR        0.0009 0.9999 2827 
mean_PPD      0.0544 0.9994 3387 
log-posterior 0.0238 1.0006 1835

For each parameter, mcse is Monte Carlo standard error, n_eff is a crude measure of effective sample size, and Rhat is the potential scale reduction factor on split chains (at convergence Rhat=1).
\end{verbatim}

\textbf{Have obviously not done this correctly as the prior should have
shifted the default model parameter for eCPR to a larger number with a
small sd}

@@

\begin{verbatim}
Warning: Removed 9 rows containing missing values (`stat_slabinterval()`).
\end{verbatim}

\begin{verbatim}
Warning: Using the `size` aesthietic with geom_segment was deprecated in ggplot2 3.4.0.
i Please use the `linewidth` aesthetic instead.
\end{verbatim}

@@

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-2.pdf}

>>>>

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
  pdfauthor={James M Brophy},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

\begin{longtable}{lrrrc}
\toprule
prior & pt\_est <- ... & llimit <- ... & ulimit <- ... & p.gt.1 <- ... \\ 
\midrule
vague & 1.32 & 0.543 & 3.21 & 0.727 \\ 
combined & 1.27 & 0.696 & 2.31 & 0.782 \\ 
enthusiastic & 1.27 & 0.708 & 2.27 & 0.79 \\ 
skeptical & 1.25 & 0.695 & 2.26 & 0.777 \\ 
\bottomrule
\end{longtable}

\hypertarget{summary-of-what-i-dont-understand}{%
\section{\texorpdfstring{\textbf{Summary of what I don't
understand}}{Summary of what I don't understand}}\label{summary-of-what-i-dont-understand}}

This makes only partial sense. 95\% CrI are narrower with informative
priors. OK, but one would think the enthusiastic prior from a trial
ARREST that was stopped prematurely for benefit 6/15 vs 0/14 successes
for eCPR vs CPR would shift the HR of 1.32 to the right but instead the
HR falls to have a 1.27. The skeptical prior from PRAGUE which was
stopped for futility (38/124 vs 24/132) does shift HR down 1.25 as
expected

So enthusiastic priors - intercept N(.03,.18), Tx N(0.429,0.495)\\
skeptical prior - intercept N(.182,.386), Tx N(0.306,0.461)

Tx for enthusiastic prior \textgreater{} Tx skeptical prior (eCPR) but
this seems to be offset by the reverse occurring for the intercept
(CPR).\\
The problem seems to be in how comparable are these previous RCTs as the
baseline success for the standard TX (CRP) is radically different 0/14
vs 24/132.

When I use a prior that uses the enthusiastic intercept prior and the
skeptical Tx prior, I get a bigger downward shift with HR 1.220 95\%
CrI(0.678 2.194) and P\textgreater1 74.5\%

Can you see any other explanation?

Do you have any other insights?

@@

@@
